# Troubleshooting {#trouble}

## Package installation

Most packages are installed from CRAN using `install.packages` or bioconductor using `BiocManager::install()`. Make sure to also install dependencies, and check to see which functions are masked by other packages if you change the order of loading libraries. 

Some require installation from github, using `devtools`. For example, `dict` installation:

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mkuhn/dict")
```

If you are using Windows and struggling with compilation, make sure `RTools` is installed properly (<https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html>). If you don't have write permissions to update packages, this may require administrator access.

## File setup

The most common error is that your file names or `PATH` is incorrect.  

## Styling

If you are using a style sheet (`style.css`), first make sure it's in the correct repository. Are default options from`Shiny` overriding the options in your style sheet? Check options for each function. For more help, see
<https://shiny.rstudio.com/articles/css.html>. 

```{r eval=FALSE}
# example fluidRow
tabItem(tabName = tabName,
        fluidRow(box(title = datasetTitle,
                     status = "primary", # these with affect styling
                     solidHeader = TRUE, # these with affect styling
                     if (is.null(des_dir) == FALSE) {includeHTML(des_dir)},
                     if (is.null(image_dir) == FALSE) {img(src = image_dir, height = 150, width = 400)}
        )),br(),
        actionLink(NS(id, "link_to_home"), "Home", icon = icon("home")),br(),br(),
        actionButton(NS(id, "load"), "Plot Graphs")
)
```

## Deployment

Simple apps can be directly deployed with `rsconnect::deployApp()`, but it is common to have issues with the first deployment *even when the app works locally*. Here are a few examples below, but your error code will help inform a starting point (in particular, is there one package highlighted as a potential issue?)

### Basics

1. Slow deployment: Uploading bundle for application and building packages will take time. Are you loading unneccessary data or packages? 
2. Memory: check `Instance Size` on shinyapps.io, this can be increased with specific pricing packages.

### Bioconductor

Because this framework uses Biocondutor packages, it is likely you need to set the repository (URL). 

```{r eval=FALSE}
# example error 1: you need to set a URL

May be unable to deploy package dependency 'AnnotationDbi'; could not
determine a repository URL for the source 'Bioconductor'.

# example error 2: the URL is set but not sufficient to compile packages, may need to try another Bioconductor version, update packages, fix compiling, turn off a firewall, or update R

Unhandled Exception: Child Task 873781155 failed: Error parsing manifest: 
Unable to determine package source for Bioconductor package
```

You can specific the bioconductor repository here (depending on updates, a very new version may be incompatible with `Shiny` briefly). You can check which repositories have been set using `getOption("repos")`.

```{r eval=FALSE}
setRepositories(addURLs = c(BioC = "https://bioconductor.org/packages/3.16/bioc")) #specify version here

# alternatively, this can be run at the top of the UI and *may* help
library(BiocManager)
options(repos = BiocManager::repositories())
```

If there is a specific package referred to, check the installation to make sure it was successful (see above). You can force a new install using `force = TRUE`, but restart R before trying again. 

If you are planning to update many packages, you can check which are valid first. 

```{r eval=FALSE}
BiocManager::valid() 
```

Still not working? A few other things to look for:

1. Package dependency mismatch between CRAN and Bioconductor (check `rsconnect::appDependencies()`)
2. Make sure you are using a released version of R (shinyapps.io doesn't support beta versions, and you may need to revert to a released version of R and update packages.)
3. Your local firewall/antivirus may be blocking deployment. 

If all else fails, try posting on the community forum for help (<https://community.rstudio.com/) and tag _Shiny_ and _shinyapp.io_ with a reproducible example (ie. _reprex_). See <https://community.rstudio.com/t/faq-tips-for-writing-r-related-questions/6824> for how to best format this. 

For reference, the following were used for our database.
```{r}
rsconnect::appDependencies()
getOption("repos")
sessionInfo()
```

Happy building!
